\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-16 paper/paper 2/dc-2015.tex

Various constraint languages exist or are being developed that support more or less of these constraint types. For our work, we focus on the following four as the ones that are most popular among data practitioners, often mentioned on mailing lists and/or being candidates or prototypes for the upcoming W3C recommendation:
\emph{Description Set Profiles (DSP)}, \emph{Resource Shapes (ReSh)}, \emph{Shape Expressions (ShEx)}, and the \emph{Web Ontology Language} \emph{(OWL)}. Despite the fact that OWL is arguably not a constraint language, it is widely used in practice as such under the closed-world and unique naming assumptions.

With its direct support of validation via \emph{SPARQL}, the \emph{SPARQL Inferencing Notation (SPIN)} is also very popular to formulate and check constraints \cite{Fuerber2010}. We consider SPIN a low-level language in contrast to the other constraint languages where specific language constructs exist to define constraints in a declarative and in comparison more intuitive way -- although SPARQL aficionados might object particularly to the latter point.

The power of SPIN is shown in 
Table \ref{tab:constraint-type-specific-expressivity}, where we list the fraction (and absolute numbers in brackets) of how many constraint types each of these languages supports \cite{BoschNolleAcarEckert2015}. 
We further see that OWL 2 is currently the most expressive high-level constraint language, at least according to the pure number of contraint types supported. This does not preclude that other constraint languages are better suited for certain applications, either because they support some types that are not supported by OWL or because the constraint representation is mroe appealing to the data practitioners -- producers as well as consumers who again might have different needs and preferences.

>>>>

Various constraint languages exist or are being developed that support more or less of these constraint types. For our work, we focus on the following four as the ones that are most popular among data practitioners, often mentioned on mailing lists and/or being candidates or prototypes for the upcoming W3C recommendation:
\emph{Description Set Profiles (DSP)}, \emph{Resource Shapes (ReSh)}, \emph{Shape Expressions (ShEx)}, and the \emph{Web Ontology Language} \emph{(OWL)}. Despite the fact that OWL is arguably not a constraint language, it is widely used in practice as such under the closed-world and unique name assumptions.

With its direct support of validation via \emph{SPARQL}, the \emph{SPARQL Inferencing Notation (SPIN)} is also very popular to formulate and check constraints \cite{Fuerber2010}. We consider SPIN as a low-level language in contrast to the other constraint languages where specific language constructs exist to define constraints in a declarative and in comparison more intuitive way -- although SPARQL aficionados might object particularly to the latter point.

The power of SPIN is shown in 
Table \ref{tab:constraint-type-specific-expressivity}, where we list the fraction (and absolute numbers in brackets) of how many constraint types each of these languages supports \cite{BoschNolleAcarEckert2015}. 
We further see that OWL 2 is currently the most expressive high-level constraint language, at least according to the pure number of constraint types supported. This does not preclude that other constraint languages are better suited for certain applications, either because they support some types that are not supported by OWL or because the constraint representation is more appealing to the data practitioners -- producers as well as consumers who again might have different needs and preferences.
